\section{Conclusion}

In this project, we explored the problem of semantic similarity and duplicate question detection using the Quora Question Pairs dataset. We implemented and compared two modern Transformer based models, BERT and SBERT, to evaluate their effectiveness and efficiency.

Our experiments showed that the standard BERT model achieved the highest overall performance metrics, including accuracy, precision, and F1-score, but suffered from significantly slower inference times. In contrast, SBERT provided a much faster alternative. Specifically, fine-tuning SBERT alongside a logistic classifier (SBERT Approach 2) yielded results that were highly competitive with BERT, while taking only a fraction of the inference time. The zero-shot cosine similarity approach (SBERT Approach 1) was the fastest method tested, but it resulted in the lowest accuracy and F1-score out of the three configurations.

Additionally, the rapid early stopping observed during our training process highlights the strong language representation capabilities already present in these pre-trained models. For future work and improvements, we could explore adding a learning rate scheduler to stabilize the rapidly fluctuating metrics during training, or investigate linear probing by only fine-tuning the classifier instead of the entire model. Finally, evaluating these models on a task that requires finding semantic matches from a large, unlabelled pool of sentences, rather than just evaluating provided positive and negative pairs, would better emphasize SBERT's efficiency advantages in real-world semantic clustering.